---
title: "Visualization"
output: 
  html_document:
    toc: true
    toc_float: true
---

This page presents visualization of the residual oil project.

```{r setup, include = FALSE}
library(tidyverse)
library(sf)
library(RColorBrewer)
library(classInt)
library(maps)
library(maptools)
library(ggpubr)
library(ggsn)
library(patchwork)
library(ggspatial)
library(leaflet)
library(BAMMtools)
library(plotly)
library(modelr)

library(RColorBrewer)
library(classInt)
library(ggsn)
library(sp)
library(spdep)
library(spgwr)
library(ggthemes)
```

## EDA and visualization

### Air pollution: PM~2.5~ and SO~2~

The data comes from [NYCCAS](https://data.cityofnewyork.us/Environment/NYCCAS-Air-Pollution-Rasters/q68s-8qxv). I am comparing data in 2011, which is the year before implementation of the Clean Heat Program, and year 2016, which is the year by which all heating oil #6 should be converted.

The pollutants of interest are PM~2.5~ and SO~2~. Here I imported the data and plotted the spatial distribution in 2011 and 2016 and the changes for these two pollutants.

```{r include = FALSE}
# data import
nyc_tract <- st_read(dsn = '/Users/zhanglvou/Desktop/GoMailman/research/number_boilers/fuel_data/data/so2_abs_lag', layer = 'model_so2_abs_lag') %>% 
  rename(
    'PM2.5_2011' = 'm__2012',
    'PM2.5_2016' = 'm__2016',
    'SO2_2011' = 'm_2_2012',
    'SO2_2016' = 'm_2_2016',
    'SO2_change' = 'so2_dff',
    'PM2.5_change' = 'pm_diff',
    'yearbuilt' = 'avg_yer'
  ) %>% 
  as_Spatial()


```

```{r warning = FALSE, include = FALSE}
# convert and join
nyc_tract@data$id <- rownames(nyc_tract@data)
gpclibPermit()
nyc_tract.fort <- broom::tidy(nyc_tract, region="id")
tract_plot <- plyr::join(nyc_tract.fort, nyc_tract@data, by="id")

```

```{r eval = FALSE}
# pm in 2011
pm_12_plot <- ggplot(tract_plot) +
  aes(long, lat, group = group, fill=cut_number(m__2012, 5)) +
  geom_polygon() +
  coord_equal() + 
  scale_fill_brewer(name="Quantile", palette="Blues") +
  labs(
    title = 'PM2.5, 2011, ug/m3',
    x = '',
    y = ''
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none')

# pm in 2016
pm_16_plot <- ggplot(tract_plot) +
  aes(long, lat, group=group, fill=cut_number(m__2016, 5)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette="Blues") +
  labs(
    title = str_c('PM2.5, 2016, ug/m3'),
    x = '',
    y = ''
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none')

# reduction in pm
pm_diff_plot <- ggplot(tract_plot) +
  aes(long, lat, group=group, fill=cut_number(pm_diff, 5)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette="Blues") +
  labs(
    title = str_c('PM2.5 (2011-2016),ug/m3'),
    x = '',
    y = ''
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none')

# relative difference
pm_rdiff_plot <- ggplot(tract_plot) +
  aes(long, lat, group=group, fill=cut_number(pm_rdff, 5)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette="Blues") +
  labs(
    title = str_c('PM2.5 (2011-2016), relative'),
    x = '',
    y = '',
    caption = 'Data source: NYCCAS'
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none')


```

```{r eval = FALSE}
# combine
(pm_12_plot + pm_16_plot)/(pm_diff_plot + pm_rdiff_plot)


```



#### PM2.5

```{r}

popup_pm <- str_c('<b>GEOID: </b>', nyc_tract$geoid, '<br>',
               '<b>PM2.5 change: </b>', round(nyc_tract$PM2.5_change, digits = 2))

pal_pm <- colorQuantile('Greens', NULL, n = 5)
leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = nyc_tract,
              fillColor = ~pal_pm(PM2.5_change),
              color = 'grey',
              weight = 1,
              fillOpacity = 0.8,
              popup = popup_pm,
              highlight = highlightOptions(
    weight = 4,
    color = "#666",
    fillOpacity = 0.3,
    bringToFront = TRUE)) %>% 
   addLegend(data = nyc_tract,
     pal = pal_pm, 
             values = ~nyc_tract$PM2.5_change, opacity = 0.7, title = NULL,
    position = "bottomright")




```

#### SO2


I made a similar plot to show the changes in SO2 from 2012 to 2016. In fact, reduction.

```{r}
popup_so2 <- str_c('<b>GEOID: </b>', nyc_tract$geoid, '<br>',
               '<b>SO2 change: </b>', round(nyc_tract$SO2_change, digits = 2))

pal_so2 <- colorQuantile('Blues', NULL, n = 5)
leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = nyc_tract,
              fillColor = ~pal_so2(SO2_change),
              color = 'grey',
              weight = 1,
              fillOpacity = 0.8,
              popup = popup_so2,
              highlight = highlightOptions(
    weight = 4,
    color = "#666",
    fillOpacity = 0.3,
    bringToFront = TRUE)) %>% 
   addLegend(data = nyc_tract,
     pal = pal_so2, 
             values = ~nyc_tract$SO2_change, opacity = 0.7, title = NULL,
    position = "bottomright")




```





These plots show that:  

*  In general, PM~2.5~ has reduced across the city, as showed in the quantile legends.  
*  in both 2011 and 2016, PM~2.5~ is very clustered in Manhattan, South Bronx and Part of Queens.  
*  Reduction of air pollution is clustered in Northern Manhattan and the Bronx.  


```{r eval = FALSE}
# pm in 2011
so2_12_plot <- ggplot(tract_plot) +
  aes(long, lat, group = group, fill=cut_number(m_2_2012, 5)) +
  geom_polygon() +
  coord_equal() + 
  scale_fill_brewer(name="Quantile", palette="Greens") +
  labs(
    title = 'SO2, 2011',
    x = '',
    y = ''
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none') +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )

# pm in 2016
so2_16_plot <- ggplot(tract_plot) +
  aes(long, lat, group=group, fill=cut_number(m_2_2016, 5)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette="Greens") +
  labs(
    title = 'SO2, 2016',
    x = '',
    y = ''
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none') +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )

so2_diff_plot <- ggplot(tract_plot) +
  aes(long, lat, group = group, fill=cut_number(so2_dff, 5)) +
  geom_polygon() +
  coord_equal() + 
  scale_fill_brewer(name="Quantile", palette="Greens") +
  labs(
    title = 'Reduction in SO2, 2011-2016',
    x = '',
    y = ''
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none') +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )


# reduction in pms
so2_rdiff_plot <- ggplot(tract_plot) +
  aes(long, lat, group=group, fill=cut_number(s2_rdff, 5)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette="Greens") +
  labs(
    title = 'Reduction in SO2, relative',
    x = '',
    y = '',
    caption = 'Data source: NYCCAS'
  ) +
  north(tract_plot) +
  theme_bw() + theme(legend.position = 'none') + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )

```

```{r eval = FALSE}
# combine
(so2_12_plot + so2_16_plot)/(so2_diff_plot + so2_rdiff_plot)
```


```{r eval = FALSE}
ro2_breaks <- getJenksBreaks(nyc_tract@data$d_ro2, 5)
ro2_change <- ggplot(tract_plot) +
  aes(long, lat, group = group, fill=cut(d_ro2, breaks = c(-6, -1, 0, 1, 6, 40, 106, 159), include.lowest = TRUE)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette = "RdBu", direction = -1) +
  labs(
    title = 'Changes in fuel oil #2'
  ) +
  north(tract_plot) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )

ro4_breaks <- getJenksBreaks(nyc_tract@data$d_ro4, 5)
ro4_change <- ggplot(tract_plot) +
  aes(long, lat, group = group, fill=cut(d_ro2, breaks = c(-31,-1, 0, 14, 46, 137, 305), include.lowest = TRUE)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette = "RdBu", direction = -1) +
  labs(
    title = 'Changes in fuel oil #4'
  ) +
  north(tract_plot) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )
  
ro6_breaks <- getJenksBreaks(nyc_tract@data$d_ro6, 5)
ro6_change <- ggplot(tract_plot) +
  aes(long, lat, group = group, fill = cut(d_ro6, breaks = c(-69, -1, 0, 6, 47, 130, 248), include.lowest = TRUE)) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette = "RdBu", direction = -1) +
  labs(
    title = 'Changes in fuel oil #6'a
  ) +
  north(tract_plot) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )


ng_breaks <- getJenksBreaks(nyc_tract@data$d_ng, 5)
ng_change <- ggplot(tract_plot) +
  aes(long, lat, group = group, fill = cut(d_ng, breaks = c(-482, -13, -1, 0,  138))) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette = "RdBu", direction = -1) +
  labs(
    title = 'Changes in natural gas'
  ) +
  north(tract_plot) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )




```

## Analysis

### Linear regression

To make things easier for my analysis, I will use the subsetted dataset (2158 obs).

```{r}
model_tract_sp <- st_read(dsn = '/Users/zhanglvou/Desktop/GoMailman/research/number_boilers/fuel_data/data/so2_abs_lag', layer = 'model_so2_abs_lag') %>% 
  filter(!is.na(so2_dff) & !is.na(s2_rdff) & !is.na(avg_yer)) %>% 
  as_Spatial(.)


```

**SO2**

```{r}
model_tract_data <- model_tract_sp@data
lm_so2 <- lm(so2_dff ~ d_ro2 + d_ro4 + d_ro6 + d_ng + d_d2 + bus + hvytrk + medtrk + car + avg_yer, data = model_tract_data)
```

check the coefficients of the linear regression model

```{r}
lm_so2 %>% 
  broom::tidy()

lm_so2 %>% 
  broom::glance()
```

R^2^ is 0.177.

plot the residuals. I expect them to be highly correlated.

```{r warning = FALSE}
model_tract_sp@data <- add_residuals(model_tract_sp@data, lm_so2)

model_tract_sp@data$id <- rownames(model_tract_sp@data)
gpclibPermit()
model_tract_sp.fort <- broom::tidy(model_tract_sp, region = "id")
model_tract_sp_plot <- plyr::join(model_tract_sp.fort, model_tract_sp@data, by = "id")


```



```{r}
ggplot(model_tract_sp_plot) +
  aes(long, lat, group = group, fill = cut(model_tract_sp_plot$resid, breaks = c(-11, -1.58, -1.01, 0, 0.945, 9))) +
  geom_polygon() +
  coord_equal() +
  scale_fill_brewer(name="Quantile", palette = "RdBu", direction = -1) +
  labs(
    title = 'Changes in natural gas'
  ) +
  north(model_tract_sp_plot) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
        )





```

It's very clustered


### Spatial Durbin

To run the durbin model, I need to exclude some missing rows in the dataset

```{r}
neighbor <- poly2nb(model_tract_sp)
list_weight <- nb2listw(neighbor)
durbin_so2 <- lagsarlm(so2_dff ~ d_ro2 + d_ro4 + d_ro6 + d_ng + d_d2, data = model_tract_sp, Durbin = TRUE, list_weight)

```

Some neighbors and list of weights 

```{r}
model_tract_sp@data$d_so2_resid <- residuals(durbin_so2)
```

```{r}
popup_dso2_resid <- str_c('<b>GEOID: </b>', model_tract_sp$geoid, '<br>',
               '<b>Residual of spatial durbin: </b>', round(model_tract_sp$d_so2_resid, digits = 2))

pal_dso2_resid <- colorQuantile('RdBu', NULL, n = 5, reverse = TRUE)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = model_tract_sp,
              fillColor = ~pal_dso2_resid(model_tract_sp$d_so2_resid),
              color = 'grey',
              weight = 1,
              fillOpacity = 0.8,
              popup = popup_dso2_resid,
              highlight = highlightOptions(
    weight = 4,
    color = "#666",
    fillOpacity = 0.3,
    bringToFront = TRUE))


```

```{r}
moran_I <- function(x, y = NULL, W){
  if(is.null(y)) y = x
  
  xp <- (x - mean(x, na.rm=T))/sd(x, na.rm=T)
  yp <- (y - mean(y, na.rm=T))/sd(y, na.rm=T)
  W[which(is.na(W))] <- 0
  n <- nrow(W)
  
  global <- (xp%*%W%*%yp)/(n - 1)
  local  <- (xp*W%*%yp)
  
  list(global = global, local  = as.numeric(local))
}

simula_moran <- function(x, y = NULL, W, nsims = 1000){
  
  if(is.null(y)) y = x
  
  n   = nrow(W)
  IDs = 1:n
  
  xp <- (x - mean(x, na.rm=T))/sd(x, na.rm=T)
  W[which(is.na(W))] <- 0
  
  global_sims = NULL
  local_sims  = matrix(NA, nrow = n, ncol=nsims)
  
  ID_sample = sample(IDs, size = n*nsims, replace = T)
  
  y_s = y[ID_sample]
  y_s = matrix(y_s, nrow = n, ncol = nsims)
  y_s <- (y_s - apply(y_s, 1, mean))/apply(y_s, 1, sd)
  
  global_sims  <- as.numeric( (xp%*%W%*%y_s)/(n - 1) )
  local_sims  <- (xp*W%*%y_s)
  
  list(global_sims = global_sims,
       local_sims  = local_sims)
}




```


### GWR

```{r}
bw_so2 <- gwr.sel(so2_dff ~ d_ro2 + d_ro4 + d_ro6 + d_ng + d_d2 + bus + hvytrk + medtrk + car + avg_yer, data = model_tract_sp, gweight = gwr.Gauss, verbose = TRUE)
gwr_so2 <- gwr(so2_dff ~ d_ro2 + d_ro4 + d_ro6 + d_ng + d_d2 + bus + hvytrk + medtrk + car + avg_yer, data = model_tract_sp, bandwidth = bw_so2, gweight = gwr.Gauss)


```

```{r}
model_tract_sp@data$gwr_r2 <- gwr_so2$SDF$localR2
model_tract_sp@data$ro2_coeff <- gwr_so2$SDF$d_ro2
model_tract_sp@data$ro4_coeff <- gwr_so2$SDF$d_ro4
model_tract_sp@data$ng_coeff <- gwr_so2$SDF$d_ng
model_tract_sp@data$d2_coeff <- gwr_so2$SDF$d_d2
model_tract_sp@data$ro4_coeff <- gwr_so2$SDF$d_ro4
model_tract_sp@data$gwr_residual <- gwr_so2$SDF$gwr.e

```



```{r}
popup_localr2 <- str_c('<b>GEOID: </b>', model_tract_sp$geoid, '<br>',
               '<b>Local R2: </b>', round(gwr_so2$SDF$localR2, digits = 2))

pal <- colorQuantile('Greens', NULL, n = 5)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = gwr_so2$SDF,
              fillColor = ~pal(gwr_so2$SDF$localR2),
              color = 'grey',
              weight = 1,
              fillOpacity = 0.8,
              popup = popup_localr2,
              highlight = highlightOptions(
    weight = 4,
    color = "#666",
    fillOpacity = 0.3,
    bringToFront = TRUE))
```

```{r}
popup_ro6 <- str_c('<b>GEOID: </b>', model_tract_sp$geoid, '<br>',
               '<b>Coefficient of #RO6 : </b>', round(gwr_so2$SDF$d_ro6, digits = 2))

bins <- getJenksBreaks(gwr_so2$SDF$d_ro6, 6)
pal_ro6 <- colorBin("RdYlBu", domain = gwr_so2$SDF$d_ro6, bins = c(-0.39763151, -0.04608157, 0,  0.12670816,  0.21932648 , 0.47350907), reverse = TRUE)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = gwr_so2$SDF,
              fillColor = ~pal_ro6(gwr_so2$SDF$d_ro6),
              color = 'grey',
              weight = 1,
              fillOpacity = 0.8,
              popup = popup_ro6,
              highlight = highlightOptions(
    weight = 4,
    color = "#666",
    fillOpacity = 0.3,
    bringToFront = TRUE)) %>% 
  addLegend(pal = pal_ro6, values = gwr_so2$SDF$d_ro6, opacity = 0.7, title = NULL,
    position = "bottomright")




```







